---
title: "Calculating Apparent Digestibility"
author: "Written by: Anıl Axel Tellbüscher"
date: "Last updated: 2025-12-31 (rebuilt `r Sys.Date()`)"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
      toc_depth: 2
bibliography: references.bib
csl: aquaculture-international.csl
vignette: >
  %\VignetteIndexEntry{digestibility}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr_setup, include = FALSE, echo=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



This short vignette provides coding examples on how to use the functions provided by the `aquacultuR` package to calculate **Apparent Digestibility Coefficients** of feeds, feed nutrients, and feed ingredients.



```{r load_packages, message=FALSE, warning=FALSE}
library(aquacultuR)
library(magrittr)
library(dplyr)
```

```{r options}
options(digits = 3)
```



# Data

*Note: We could not find any published dataset on digestibility trials in data journals or Open Access repositories (Zenodo, Mendeley Data). We would thus appreciate data contributions from the aquaculture community!* 

The *digestdm* dataset contains two fictious diets that differ in their dry matter content and the mass fraction of marker that was found in the feces of an imaginary fish species.

```{r}
digestdm
```

The *digestnut*  dataset contains the same two fictious feeds but the dataset is enriched by fictious total nitrogen contents from which the crude protein content of the samples can be estimated.

```{r}
digestnut
```

While the example data for the dry matter and specific nutrient digestibility calculations provided are constructed, the data for the calculation of ingredient digestibility originates from @Bureau1999a and @Bureau2006a. It contains all required data for a blood meal that was tested for apparent digestibility in the scope of the cited studies.

```{r}
digestingr
```



# Dry matter digestibility

Based on the *digestdm* dataset, the **Apparent Digestibility Coefficient** for the dry matter fraction of a feed ($\text{ADC}_{\text{DM}}$) can immediately be calculated using the `adc_dm()` function.

```{r calculate_adc_dm_short}
digestdm %>%
  group_by(diet) %>%
  summarise(`ADC DM` = adc_dm(std_diet = std_feed, std_feces = std_feces))
```

Note that the function default for the dry matter content of the feed is set to 1. This can be changed with the `dm` argument of the `adc_dm()` function. The dry matter content of the feces cannot be adjusted.

```{r calculate_adc_dm_additional_arguments}
digestdm %>%
  group_by(diet) %>%
  summarise(`ADC DM` = adc_dm(
    dm = dm,
    std_diet = std_feed,
    std_feces = std_feces
  ))
```

Applied on a data frame or tibble, the code outputs the $\text{ADC}_{\text{DM}}$ for each diet.





# Nutrient digestibility

The **Apparent Digestibility Coefficient** for specific nutrients in a feed ($\text{ADC}_{\text{nut}}$) can be calculated analogously with the `adc_nut()` function. Here, additional function arguments are the mass fractions of the nutrient of interest in the feed and the feces in gram per gram.

```{r calculate_adc_nut}
digestnut %>%
  group_by(diet) %>%
  summarise(
    `ADC CP` = adc_nut(
      std_diet = std_feed,
      std_feces = std_feces,
      nut_diet = N_feed,
      nut_feces = N_feces
    )
  )
```





# Ingredient digestibility

Lastly, the **Apparent Digestibility Coefficient** can also be calculated for the dry matter fraction of a specific ingredient in a feed ($\text{ADC}_{\text{ingr}}$). This is possible with the `adc_ingr()` function that 

```{r calculate_adc_ingr_short}
digestingr %>%
  summarise(
    `ADC Ingredient` = adc_ingr(
      adc_test = adc_test,
      adc_ref = adc_reference,
      nut_ref = CP_reference,
      nut_ingr = CP_ingr
    )
  )
```

Same as `adc_dm()`, the function assumes that the dry matter of the reference feed and the feed ingredient is 1 gram per gram by default. To change this, the function arguments `dm_ref` and `dm_ingr` can be used. In addition, the mass fraction of the feed ingredient into the test diet can be set by using the function argument `incl_ingr`. The default is 0.3 gram per gram.

```{r calculate_adc_ingr_additional_arguments}
digestingr %>%
  summarise(
    `ADC Ingredient` = adc_ingr(
      adc_test = adc_test,
      adc_ref = adc_reference,
      nut_ref = CP_reference,
      nut_ingr = CP_ingr,
      dm_ref = dm_reference,
      dm_ingr = dm_ingr
    )
  )
```

The difference from the result stated in @Bureau2006a ($\text{ADC} =  87.5\%$) originates from rounding of the result. 





# Session info

```{r show_session_info}
sessionInfo()
```





# References

