---
title: "aquacultuR"
author: "Anıl Axel Tellbüscher"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: references.bib
csl: aquaculture-international.csl
vignette: >
  %\VignetteIndexEntry{aquacultuR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(aquacultuR)
library(magrittr)
library(dplyr)
```

This package consist of key formulas and functions used in aquaculture practices,these formulas help quantify feed utilization, growth performance,nutrient retention and digestiblity, This vignette provides coding examples that are relatable and simple to follow.

The following metrics are demonstrated:
1. Digestibility metrics: adc_dm(), adc_ingr(), adc_nut()
2. Growth metrics: ag(), agr(), sgr(), tgc(), rgr()
3. Feed metrics: fcr(), fce(), ner(),retention()
4. Biomass metrics: gbw(), mbw(), rfr()
5. Data handling: rearing and weight sample datasets.



## Example data

The data we will assess originates from one out of two experiments during with Atlantic salmon (*Salmo salar*) has been exposed to different levels of dissolved oxygen saturation. The original data was published by @Liland2024a in form of a typical Excel workbook and can be found together with additional information in the dedicated publication. The respective data has been tidied by converting the double-row into single-row column names.


## Growth

Let us start with growth data. 

```{r}
glimpse(samplings)
```



## References:

## Session info
```{r}
sessionInfo()
```

