---
title: "aquacultuR"
author: "Anıl Axel Tellbüscher"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{aquacultuR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

When working in aquaculture, you will be interested in calculating some key parameters to quantify

- growth performance
- feed conversion
- feed digestibility

The aquacultuR package provides functions to calculate key parameters used in aquaculture practices. This vignette demonstrates exemplarily how to use those functions by applying them to sample datasets from aquaculture experiments in a series of coding examples.

The use of the following functions is demonstrated:
1. Digestibility metrics: `adc_dm()`, `adc_ingr()`, `adc_nut()`
2. Growth metrics: `ag()`, `agr()`, `sgr()`, `tgc()`, `rgr()`
3. Feed metrics: `fcr()`, `fce()`, `ner()`, `retention()`
4. Biomass metrics: `gbw()`, `mbw()`, `rfr()`
5. Data handling: `rearing` and `weight` sample dataset.


## Load package and dependencies

```{r}
library(aquacultuR)
library(magrittr)
library(dplyr)
```



## Load example data



```{r}
data("rearing")
str(rearing)
```


```{r}
data("weight")
str(weight)
```


## Data preparation

We now merge the rearing and weight dataset.

```{r}
sample_data <- rearing %>% 
  left_join(weight, join_by(tank)) 
```

Now, let's calculate some key metrics!

```{r}
metrics_table <- sample_data %>% 
  mutate(
    ag = ag(bodyweight_g_start, bodyweight_g_end),
    agr = agr(bodyweight_g_start, bodyweight_g_end, duration),
    fcr = fcr(bodyweight_g_start, bodyweight_g_end, feed_g),
    sgr = sgr(bodyweight_g_start, bodyweight_g_end, duration),
    tgc = tgc(bodyweight_g_start, bodyweight_g_end, duration, sum(temp_C)),
    fce = fce(bodyweight_g_start, bodyweight_g_end, feed_g),
    gbw = gbw(bodyweight_g_start, bodyweight_g_end,n),
    mbw = mbw(bodyweight_g_start, bodyweight_g_end),
    rfr = rfr(fish_survived = n, fish_stocked = initial_n),
    rgr = rgr(bodyweight_g_start, bodyweight_g_end, duration) 
  )

head(metrics_table)
```





## Digestiblity calculations:

### 1. Dry matter ADC:
Measures the proportion of dry matter in feed that is apparently digested and absorbed by the animal. Calculated using an external marker (e.g., chromic oxide) in feces and feed. Indicates how well an animal utilizes the entire feed. Poor digestibility of dry matter suggests that a significant portion of feed is wasted, reducing nutritional efficiency and increasing environmental waste.

```{r}
#Function
adc_dm <- 1 - (es_diet / es_feces)

#Sample
adc_dm(es_diet = 0.01, es_feces = 0.04) 
```


### 2. Ingredient-level ADC:
Assesses the digestibility of a nutrient (e.g., protein, lipid) from a specific ingredient (e.g., soybean meal, fishmeal). Requires the ingredient to be tested in a reference diet. Helps evaluate the nutritional value of individual ingredients. Essential for selecting cost-effective, sustainable feed components in diet formulation.


```{r}
#Function
adc_ingr <- ( adc_test + ( ((1-incl_ingr) * nut_ref) / (incl_ingr * nut_ingr) ) * (adc_test - adc_ref) )
#Sample
adc_ingr(
  adc_test = 0.80,
  adc_ref = 0.75,
  nut_ref = 0.20,
  nut_ingr = 0.35,
  incl_ingr = 0.30
)
```


### 3. Nutrient-level ADC:
Calculates how much of a specific nutrient (e.g., protein, fat, energy) in the complete diet is absorbed by the animal.Critical for understanding nutrient utilization and for balancing diets to meet species-specific nutritional needs while minimizing nutrient waste.

```{r}
#Function
adc_nut <- (1 - (es_diet / es_feces) * (nut_feces / nut_diet))
#sample
adc_nut(
  nutrient_diet = 0.35,
  nutrient_feces = 0.10,
  marker_diet = 0.01,
  marker_feces = 0.03
)
```





## Nutrient retention
The Nutrient Retention is the absolute change in the mass of a nutrient in the tissue of an organism in relation to the biomass increase.

```{r}
#Function
retention <- m_end * nut_end - m_start * nut_start
#Sample
retention(
  m_start = 2.5,      # total nutrient gain in biomass (e.g., g protein)
  m_end = 1.0,          # total nutrient fed (e.g., g protein fed)
  nut_end=12,
  nut_start=20
)
# 
```





## Nutrient Efficiency Ratio (NER)
The Nutrient Efficiency Ratio indicates how the inclusion of a specific dietary nutrient is translated into growth. It is an important metric for optimizing nutrient levels in feed.

$$
NER = Nutrient Retained / Nutrient Consumed
$$



```{r}
##Have to write for 2 scenarios
ner(
  energy_retained = 5,   # MJ/kg
  energy_intake = 20     # MJ/kg
)
```





## Growth and feed efficiency

### Absolute Growth
Absolute growth represents the total increase in body weight over a certain time period. It is a straightforward measure calculated by subtracting initial weight from final weight. While it does not account for time or size differences, it gives a direct measure of how much the animal grew. This is particularly useful in short-term trials or in assessing harvest weight in aquaculture systems.

```{r}
ag(ibw = 50, fbw = 120)
```


### Absolute Growth Rate
This is a time-adjusted version of absolute growth, representing the daily growth rate in units like grams per day. It gives a clearer view of growth efficiency over time and is helpful for comparing growth performance across multiple feeding periods or experimental groups.

```{r}
agr(ibw = 50, fbw = 120, day = 30)
```


### Feed Conversion Ratio (FCR)
FCR is the amount of feed needed to produce one unit of body weight gain. A lower FCR means more efficient feed use. It is one of the most critical metrics in animal production because it directly impacts feed cost and environmental impact. High FCR values indicate poor efficiency, requiring more feed per unit of gain.

```{r}
fcr(ibw = 50, fbw = 120, feed = 100)
```


### Feed Conversion Efficiency (FCE)
FCE is the inverse of FCR and indicates how much body mass is gained per unit of feed. Higher FCE values are desirable as they reflect more efficient conversion of feed into growth. This is a direct measure of feeding success and economic efficiency.

```{r}
fce(ibw = 50, fbw = 120, feed = 100)
```


### Specific Growth RAte (SGR)
Specific growth rate expresses the relative daily growth of an animal using a logarithmic scale. This accounts for exponential growth and allows for comparison between animals of different sizes. It's a standard metric in aquaculture and animal nutrition research due to its precision and relevance across species.

```{r}
sgr(ibw = 50, fbw = 120, days = 30)
```



### Thermal GRowth Coefficient (TGC)

TGC adjusts growth data by factoring in the influence of temperature on metabolic rate and growth. This is especially important in aquaculture, where temperature plays a key role in feeding behavior and nutrient utilization. It allows comparison of growth rates across systems with different temperatures, supporting better management and forecasting

```{r}
tgc(ibw = 50, fbw = 120, days = 30, temp = 22)
```



### Relative Growth Rate

RGR is a measure of the percentage increase in weight over time relative to the initial weight. It is simpler than SGR and does not use logarithms. Though it’s less precise for comparing across sizes, it is still useful for evaluating growth trends.

```{r}
rgr(ibw = 50, fbw = 120, days = 30)
```



### Geometric Bodyweight (GBW)

Geometric Body Weight is a calculated average used in growth studies to represent body weight when growth occurs exponentially rather than linearly. It is typically computed as the square root of the product of the initial and final weights:

$$
GBW = √(Initial Weight × final Weight).
$$

This method provides a more realistic average body weight during the experimental period, especially when animals grow rapidly. It is often used in calculating temperature-adjusted growth metrics like the Thermal Growth Coefficient (TGC). By using the geometric mean, researchers avoid biases that could arise from simply averaging initial and final weights in cases of rapid or uneven growth.

```{r}
gbw(ibw = 50, fbw = 120, n = 20)
```

### Relative Feeding Rate (RFR)

The Relative Feeding Rate, usually abbreviated as Feeding Rate expresses the daily feed intake as a percentage of the animal’s body weight. It is calculated using the formula:

$$
RFR = (Feed Intake / Body Weight)
$$

This metric helps to normalize feeding across animals of different sizes or across groups, making it easier to compare intake levels and feeding strategies. RFR is especially important in aquaculture and animal husbandry where feeding must be adjusted to the metabolic needs of animals, which scale with body weight. Optimizing RFR ensures adequate nutrition without overfeeding, which supports healthy growth, improves feed efficiency, and reduces waste and cost.



## Summary of metrics:
| Function      | Metric                        | Description                            |
| ------------- | ----------------------------- | -------------------------------------- |
| `adc_dm()`    | ADC of Dry Matter             | Digestibility based on external marker |
| `adc_ingr()`  | ADC of Ingredient Nutrient    | Digestibility of a feed ingredient     |
| `adc_nut()`   | ADC of Nutrient               | Digestibility of a specific nutrient   |
| `ag()`        | Absolute Growth               | Weight gain (end - start)              |
| `agr()`       | Absolute Growth Rate          | Growth over time                       |
| `fcr()`       | Feed Conversion Ratio         | Feed given per unit weight gain        |
| `fce()`       | Feed Conversion Efficiency    | Weight gain per unit feed              |
| `sgr()`       | Specific Growth Rate          | Relative growth per day                |
| `tgc()`       | Thermal Growth Coefficient    | Growth corrected for temperature       |
| `rgr()`       | Relative Growth Rate          | Percent growth per unit time           |
| `gbw()`       | Group Biomass Gain            | Total biomass gain in group            |
| `mbw()`       | Mean Biomass Gain             | Mean biomass change                    |
| `rfr()`       | Relative Fish Recovery        | Survival or harvest yield ratio        |
| `ner()`       | Net Energy Retention          | Energy retained vs. consumed           |
| `retention()` | Nutrient Retention Efficiency | Nutrient retained vs. intake           |



## References:



