---
title: "aquacultuR"
author: "Anıl Axel Tellbüscher"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{aquacultuR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(aquacultuR)
library(magrittr)
library(dplyr)
```

This package consist of key formulas and functions used in aquaculture practices,these formulas help quantify feed utilization, growth performance,nutrient retention and digestiblity, This vignette provides coding examples that are relatable and simple to follow. AquacultuR is a mechanisitic and matghematical package which is applicable to various zootechnical parameters/ metrics. 

The following metrics are demonstrated:
1. Digestibility metrics: adc_dm(), adc_ingr(), adc_nut()
2. Growth metrics: ag(), agr(), sgr(), tgc(), rgr()
3. Feed metrics: fcr(), fce(), ner(),retention()
4. Biomass metrics: gbw(), mbw(), rfr()
5. Data handling: rearing and weight sample dataset.

## Load example data

we start by merging the rearing and weight dataset and applying them in aquaculture formulas.

```{r}
data("rearing")
str(rearing)

data("weight")
str(weight)

rearing %>% 
  left_join(weight, join_by(tank)) %>% 
  mutate(
    ag = ag(bodyweight_g_start, bodyweight_g_end),
    agr = agr(bodyweight_g_start, bodyweight_g_end, duration),
    fcr = fcr(bodyweight_g_start, bodyweight_g_end, feed_g),
    sgr = sgr(bodyweight_g_start, bodyweight_g_end, duration),
    tgc = tgc(bodyweight_g_start, bodyweight_g_end, duration, sum(temp_C)),
    fce = fce(bodyweight_g_start, bodyweight_g_end, feed_g),
    gbw = gbw(bodyweight_g_start, bodyweight_g_end,n),
    mbw= mbw(bodyweight_g_start, bodyweight_g_end),
    rfr = rfr ( fish_survived=n, fish_stocked = initial_n),
    rgr = rgr( bodyweight_g_start, bodyweight_g_end, duration) 
  ) -> metrics_table
head(metrics_table)
```
## Digestiblity calculations:

#1. Dry matter ADC:

# Measures the proportion of dry matter in feed that is apparently digested and absorbed by the animal. Calculated using an external marker (e.g., chromic oxide) in feces and feed. Indicates how well an animal utilizes the entire feed. Poor digestibility of dry matter suggests that a significant portion of feed is wasted, reducing nutritional efficiency and increasing environmental waste.

```{r}
#Function
adc_dm <- 1 - (es_diet / es_feces)
#Sample
adc_dm(es_diet = 0.01, es_feces = 0.04) 

```
#2. Ingredient-level ADC:

#Assesses the digestibility of a nutrient (e.g., protein, lipid) from a specific ingredient (e.g., soybean meal, fishmeal). Requires the ingredient to be tested in a reference diet. Helps evaluate the nutritional value of individual ingredients. Essential for selecting cost-effective, sustainable feed components in diet formulation.


```{r}
#Function
adc_ingr <- ( adc_test + ( ((1-incl_ingr) * nut_ref) / (incl_ingr * nut_ingr) ) * (adc_test - adc_ref) )
#Sample
adc_ingr(
  adc_test = 0.80,
  adc_ref = 0.75,
  nut_ref = 0.20,
  nut_ingr = 0.35,
  incl_ingr = 0.30
)
```
#3. Nutrient-level ADC:

#Calculates how much of a specific nutrient (e.g., protein, fat, energy) in the complete diet is absorbed by the animal.Critical for understanding nutrient utilization and for balancing diets to meet species-specific nutritional needs while minimizing nutrient waste.

```{r}
#Function
adc_nut <- (1 - (es_diet / es_feces) * (nut_feces / nut_diet))
#sample
adc_nut(
  nutrient_diet = 0.35,
  nutrient_feces = 0.10,
  marker_diet = 0.01,
  marker_feces = 0.03
)
```
##Nutrient retention

# Percentage of a consumed nutrient retained in the body:
# Retention (%) = 100 × (Final Body Nutrient – Initial) / Nutrient Intake.
#Measures how well the animal utilizes dietary nutrients for tissue growth. High retention is good for cost and environmental reasons (less waste).

```{r}
#Function
retention <- m_end * nut_end - m_start * nut_start
#Sample
retention(
  m_start = 2.5,      # total nutrient gain in biomass (e.g., g protein)
  m_end = 1.0,          # total nutrient fed (e.g., g protein fed)
  nut_end=12,
  nut_start=20
)
# 
```
##Net energy retention (NER)

#Proportion of consumed energy retained in body tissues:
#NER = Energy Retained / Energy Consumed.
#Indicates how efficiently dietary energy is used for growth vs. being lost in metabolism. Important for optimizing energy levels in feed.

```{r}
##Have to write for 2 scenarios
ner(
  energy_retained = 5,   # MJ/kg
  energy_intake = 20     # MJ/kg
)
```
## Growth and feed efficiency
#Ag

#Absolute growth represents the total increase in body weight over a certain time period. It is a straightforward measure calculated by subtracting initial weight from final weight. While it does not account for time or size differences, it gives a direct measure of how much the animal grew. This is particularly useful in short-term trials or in assessing harvest weight in aquaculture systems.

```{r}
ag(initial = 50, final = 120)
```

#agr

#This is a time-adjusted version of absolute growth, representing the daily growth rate in units like grams per day. It gives a clearer view of growth efficiency over time and is helpful for comparing growth performance across multiple feeding periods or experimental groups.

```{r}
ag(initial = 50, final = 120, days=30)
```

#fcr

#FCR is the amount of feed needed to produce one unit of body weight gain. A lower FCR means more efficient feed use. It is one of the most critical metrics in animal production because it directly impacts feed cost and environmental impact. High FCR values indicate poor efficiency, requiring more feed per unit of gain.

```{r}
fcr(initial = 50, final = 120, feed = 100)
```

#fce

#FCE is the inverse of FCR and indicates how much body mass is gained per unit of feed. Higher FCE values are desirable as they reflect more efficient conversion of feed into growth. This is a direct measure of feeding success and economic efficiency.

```{r}
fce(initial = 50, final = 120, feed = 100)
```

#sgr

#Specific growth rate expresses the relative daily growth of an animal using a logarithmic scale. This accounts for exponential growth and allows for comparison between animals of different sizes. It's a standard metric in aquaculture and animal nutrition research due to its precision and relevance across species.

```{r}
sgr(initial = 50, final = 120, days = 30)
```

#tgc

#TGC adjusts growth data by factoring in the influence of temperature on metabolic rate and growth.This is especially important in aquaculture, where temperature plays a key role in feeding behavior and nutrient utilization. It allows comparison of growth rates across systems with different temperatures, supporting better management and forecasting

```{r}
tgc(initial = 50, final = 120, days = 30, temp = 22)
```

#rgr

#RGR is a measure of the percentage increase in weight over time relative to the initial weight. It is simpler than SGR and does not use logarithms. Though it’s less precise for comparing across sizes, it is still useful for evaluating growth trends.

```{r}
rgr(initial = 50, final = 120, days = 30)
```

#gbw

#Geometric Body Weight (GBW) is a calculated average used in growth studies to represent body weight when growth occurs exponentially rather than linearly. It is typically computed as the square root of the product of the initial and final weights:
#GBW = √(Initial Weight × Final Weight).
#This method provides a more realistic average body weight during the experimental period, especially when animals grow rapidly. It is often used in calculating temperature-adjusted growth metrics like the Thermal Growth Coefficient (TGC). By using the geometric mean, researchers avoid biases that could arise from simply averaging initial and final weights in cases of rapid or uneven growth.

```{r}
gbw(initial = 50, final = 120, n = 20)
```

#rfr

#Relative Feeding Rate (RFR) expresses the daily feed intake as a percentage of the animal’s body weight. It is calculated using the formula:
#RFR (%) = (Feed Intake / Body Weight) × 100.
#This metric helps to normalize feeding across animals of different sizes or across groups, making it easier to compare intake levels and feeding strategies. RFR is especially important in aquaculture and animal husbandry where feeding must be adjusted to the metabolic needs of animals, which scale with body weight. Optimizing RFR ensures adequate nutrition without overfeeding, which supports healthy growth, improves feed efficiency, and reduces waste and cost.

```{r}
rfr(fish_survived = 18, fish_stocked = 20)
```

## Summary of metrics:
| Function      | Metric                        | Description                            |
| ------------- | ----------------------------- | -------------------------------------- |
| `adc_dm()`    | ADC of Dry Matter             | Digestibility based on external marker |
| `adc_ingr()`  | ADC of Ingredient Nutrient    | Digestibility of a feed ingredient     |
| `adc_nut()`   | ADC of Nutrient               | Digestibility of a specific nutrient   |
| `ag()`        | Absolute Growth               | Weight gain (end - start)              |
| `agr()`       | Absolute Growth Rate          | Growth over time                       |
| `fcr()`       | Feed Conversion Ratio         | Feed given per unit weight gain        |
| `fce()`       | Feed Conversion Efficiency    | Weight gain per unit feed              |
| `sgr()`       | Specific Growth Rate          | Relative growth per day                |
| `tgc()`       | Thermal Growth Coefficient    | Growth corrected for temperature       |
| `rgr()`       | Relative Growth Rate          | Percent growth per unit time           |
| `gbw()`       | Group Biomass Gain            | Total biomass gain in group            |
| `mbw()`       | Mean Biomass Gain             | Mean biomass change                    |
| `rfr()`       | Relative Fish Recovery        | Survival or harvest yield ratio        |
| `ner()`       | Net Energy Retention          | Energy retained vs. consumed           |
| `retention()` | Nutrient Retention Efficiency | Nutrient retained vs. intake           |

## Conclusion

These metrics are essential for assessing performance, efficeiency and digetsiblity in aquuaculture practices and growth trails based on nutrition. The aquacultuR package provides standardized functions to streamline such calculations across studies.

## References:

## Session info
```{r}
sessionInfo()
```

